Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2023-01-17 supplement.tex
update best fit yaml models
Update supplement.tex

Our models also included recent events to account for known migrations,
admixtures, and growths and declines in effective population sizes. Many of
these parameters were fixed based on previous historical, genetic, or
anthropological research, namely
\begin{itemize}
    \item Neanderthal introgression to Eurasian branch: 45ka
    \item Mende population expansion and Gumuz population decline: 5ka \citep{Gopalan2022-pw}
    \item East African pastoralist to South African admixture: 2ka \citep{Uren2016-nn}
    \item European to South African admixture: 10 generations ago
    \item South African population decline following colonial admixture: 9 generations ago
\end{itemize}
The total number of parameters that were
ultimately inferred were 21-31, depending on the complexity
of the model.

@@

In our preliminary models, the unconstrained inferred split times between East
African populations and Europeans was inferred to have occurred around $40$ka,
and the inferred split between East and West African populations occurred about
$10$ka earlier. These recent split dates are at odds with paleoanthropological 
evidence for an earlier Out-of-Africa (OOA) expansion
\citep[e.g.][]{Fu2014-am,Hublin2020-nk,Hajdinjak2021-jo}, with the estimated date
for Neanderthal admixture into Eurasian populations
\citep[e.g.][]{Reilly2022-ym}, and with previous genetics-based dating of
splits between African and Eurasian populations
\citep[e.g.]{Gutenkunst2009-ff,Jouganous2017-pq,Kamm2020-vn}. The parameter
uncertainty of our inferred divergence times (which accounts for the
non-independence of sites) does not account for these differences: in the
high-likelihood models, the $95\%$ confidence intervals for the East
African-European split was roughly $38 \pm 5$ka, and for the East/West African
split was $45 \pm 10$ka, with the uncertainties reflecting 2 standard
deviations according to bootstrap of genomic regions. We therefore expect these
discrepancies to result from model misspecification in the details of the
divergence between the ancestors of OOA and East African populations.

@@

\section{Model validation results}\label{sec:supp-results}
\label{sec:validation_results}
\subsection{Model fit using alternative datasets}
\label{sec:supp-alt-data}

>>>>

Our models also included recent events to account for known migrations,
admixtures, and growths and declines in effective population sizes. Many of
these parameters were fixed based on previous historical, genetic, or
anthropological research, namely
\begin{itemize}
    \item Western African vs. Eastern African divergence: 60ka \citep{Pagani2015-pz}
%Add Niang et al. "The Middle Stone Age occupations of Tiemassa, coastal West Africa, between 62 and 25 thousand years ago" Journal of Archaeological Science, 2020 as citation
%Add 
    \item Eastern Africa vs. British divergence [i.e. OOA]: 50ka
    \item Neanderthal introgression to Eurasian branch: 45ka \citep{{Reilly2022-ym}
    \item Mende population expansion: 5ka 
    \item Gumuz population decline: 5ka
    \item East African pastoralist to South African admixture: 2ka 
    \item European to South African admixture: 10 generations ago \citep{Uren2016-nn}
    \item South African population decline following colonial admixture: 9 generations ago \citep{Van_Eeden2022-od}
\end{itemize}
The total number of parameters that were ultimately inferred were 21-31, depending on the complexity
of the model.

@@

In our preliminary models, the unconstrained inferred split times between Eastern
African populations and Europeans was inferred to have occurred around $40$ka,
and the inferred split between Eastern and Western African populations occurred about
$10$ka earlier. These recent split dates are at odds with paleoanthropological 
evidence for an earlier Out-of-Africa (OOA) expansion
\citep[e.g.][]{Fu2014-am,Hublin2020-nk,Hajdinjak2021-jo}, with the estimated date
for Neanderthal admixture into Eurasian populations
\citep[e.g.][]{Reilly2022-ym}, and with previous genetics-based dating of
splits between African and Eurasian populations
\citep[e.g.]{Gutenkunst2009-ff,Jouganous2017-pq,Kamm2020-vn}. The parameter
uncertainty of our inferred divergence times (which accounts for the
non-independence of sites) does not account for these differences: in the
high-likelihood models, the $95\%$ confidence intervals for the Eastern
African-European split was roughly $38 \pm 5$ka, and for the Eastern/Western African
split was $45 \pm 10$ka, with the uncertainties reflecting 2 standard
deviations according to bootstrap of genomic regions. We therefore expect these
discrepancies to result from model misspecification in the details of the
divergence between the ancestors of OOA and Eastern African populations.

@@

\section{Model validation results}\label{sec:supp-results}
\label{sec:validation_results}
\subsection{Model fits using alternative datasets}
\label{sec:supp-alt-data}
